Lifting CDCL to Numerical Domains

### 2016-10-11 contributions.tex
example added

>>>>

\item We present an abstract model search algorithm that strictly 
generalizes BCP procedure to compute fix-point over template polyhedra 
domain.  Our abstract conflict analysis algorithm present a non-trivial 
lifting of the Unique Implication Point (UIP) procedure to abstract 
lattices in order to learn a generalized reason for conflicts.

### 2016-10-11 example.tex
example added

\section{Motivating Example}
%
\begin{figure}[t]
\scriptsize
\begin{tabular}{l|l|l}
\hline
C program & Control-Flow Graph & Decision Procedure Statistics\\
\hline
\begin{lstlisting}[mathescape=true,language=C]
void main() {
 int x,result1=0,result2=0;
 assume(x>=1 && x<= N);
 result1 = 2*x; 
 int state = 1;
 int y=x;
 while(state!=0) {
  switch(state) {
   case 0: state = 1; break;
   case 1: state = 2; break;
   case 2: x--; y++;
   if(x == 0) state = 0;
   else state = 1; break;
  } 
 }
 result2 = y;
 assert(result1 == result2);
}
\end{lstlisting}
&
\begin{minipage}{4.40cm}
\centering
\vspace*{0.3cm}
\scalebox{.45}{\import{figures/}{example.pspdftex}}
%\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
\end{minipage}
&
\begin{lstlisting}[mathescape=true,language=C]
                 SAT solver  ACDCL 
                 ----------- ----- 
restarts           1          0
conflicts  :       5          2
decisions  :       36         2 
propagations :     1365       240  
conflict literals  65         4   
\end{lstlisting}
\\
\hline
\end{tabular}
\caption{C Program, CFG and statistics for decision procedure}
\label{example}
\end{figure}
Let us consider a simple program in Figure~\ref{example} that checks 
the equivalence of a state-machine based interger multiplication against  
a specification that uses multiplication operator (*).  The assertion 
simply compares the result of two multiplication for an input $x$.  To 
avoid overflow error, we assume $x$ has an upper bound $N$.  The 
control flow graph is shown in the middle column of Figure~\ref{example}.

We analyze the program using three different analyzers -- {\em bounded 
model checking (BMC)}, {\em abstract interpretation} and {\em ACDCL}.    
For $N=2$ and unwind limit of $bound=4$ for the {\em while} loop, the 
right column shows the statistics from MiniSAT~\cite{minisat} solver 
for analysis using BMC.  The SAT solver makes 36 decisions, 1365 boolean 
constraint propagations and performs 5 conflict analysis with 65 conflict 
literals.  Moreover, the solver performs a restart, presumably to recover 
from bad decisions.

On the other hand, abstract interpretation using forward analysis 
is too imprecise to verify the safety of the above program using 
interval domain.  The imprecision is due to control-flow join at 
node $n1$.  State-of-the-art abstract interpretation tool, Astr{\'e}e, 
requires external hints, provided by manually annotating the code 
with partition directives at $n1$, to prove safety.  The partition 
directive helps the tool to analyze each program execution paths separately.  
To do so, Astr{\'e}e requires a trace partitioning domain along with interval 
to determine safety.  Consequently, the analysis becomes very expensive 
for a large value of $N$.  In general, the imprecision is either intended 
  by the tool because such high precision analysis is normally not required for runtime error analysis or the imprecision is unavoidable due to the complexity 
of the application under analysis.

Figure~\ref{} shows an analysis using ACDCL with interval domain.  The 
analysis associates an interval for each control location and variable, 
as shown in the abstract conflict graph.

The result of the analysis is shown in right side of Figure~\ref{example}.  
Compared to a SAT solver, ACDCL makes only a single decisions which lead 
to 240 interval constraint propagations, and learns a single conflict clause 
involving only one literal.

\begin{figure}[htp]
\centering
\vspace*{0.3cm}
\scalebox{.45}{\import{figures/}{cfg-conflict.pspdftex}}
\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
\end{figure}

\Omit {
\lstdefinestyle{mystyle}{
    numbers=left,                    
    numbersep=5pt,                  
    tabsize=2,
    mathescape=true,
    language=C
}
\lstset{style=mystyle}
}
Let us consider a simple safe program $P$ as shown in Figure~\ref{fig:filter}A.
Astr{\'e}e fails to verify the safety using interval and trace-partition domain
due to control-flow join at location $n5$.  Astr{\'e}e requires external hints, provided by 
manually annotating the code with partition directives at $n1$, to prove safety.  
In general, the imprecision is either intended by the tool because such high precision analysis is 
normally not required for runtime error analysis or the imprecision is unavoidable 
due to the complexity of the application under analysis.

On the other hand, ACDCL uses decision and learning to generate the proof 
using simple interval domain. The analysis associates an interval with each location and variable.  
Assuming $bool: c = [0,1]$, ACDCL performs backward propagation 
starting from $n6$ by computing the weakest precondition for every statement 
in $P$ and immediately infers that (\(n4: x = \bot)\), thus proving safety as shown in 
red color in figure~\ref{fig:filter}A.  Now, let us consider program $P'$ in
Figure~\ref{fig:filter}B. Assuming $char: c$, ACDCL performs both forward and
backward propagations (shown in red color in Figure~\ref{fig:filter}B).  
The analysis then perform sequence of decisions starting with $c = [0,255]$ and reaches a 
conflict when the decision is (\(n1:c == [0,0]\)) which is connected to $(n6: \bot)$ 
and suffices to prove safety.  The deductions made during fixed point
iteration are represented by abstract conflict graph shown in Figure~\ref{fig:filter}C. 
Similar to conflict analysis phase in SAT solvers, ACDCL learns (\(n1: c = [1,255]\)), 
that is all error traces must satisfy (\(c \neq 0\)) at n1.  The analysis backtracks 
discarding all assumptions. Interval analysis is run with the learnt constraint and 
can prove safety.  Thus, decisions and clause learning are used to avoid case based reasoning. 
The advantage of ACDCL over propositional SAT solver is that the decision heuristics in 
ACDCL can exploit the program structure by making decisions on interesting program 
variables (for example variables in conditional branches or loop variables). Experimental 
evidence shows that this leads to significantly less number of decisions compared 
to the propositional solver. Compared to classical abstract interpretation 
based tools, ACDCL automatically performs program and property driven trace partitioning 
using decision and clause learning to generate proofs using simpler domain -- thus, it is more 
precise than classical abstract interpretation.

>>>>

\section{Motivating Example}
%
\begin{figure}[t]
\scriptsize
\begin{tabular}{l|l|l}
\hline
C program & Control-Flow Graph & Abstract Conflict Graph \\
\hline
\begin{lstlisting}[mathescape=true,language=C]
void main() {
int x,sepc=0,impl=0;
assume(x>=1 && x<=N);
spec = 2*x; 
int state = 1;
int y=x;
while(state!=0) {
 switch(state) {
  case 0: state = 1; 
  break;
  case 1: state = 2; 
  break;
  case 2: x--; y++;
  if(x == 0) state=0;
  else state=1;break;
 } 
}
impl = y;
assert(spec == impl);
}
\end{lstlisting}
&
\begin{minipage}{4.40cm}
\centering
\vspace*{0.3cm}
\scalebox{.45}{\import{figures/}{example.pspdftex}}
%\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
\end{minipage}
&
\begin{minipage}{4.40cm}
\centering
\vspace*{0.3cm}
\scalebox{.45}{\import{figures/}{acdl_run.pspdftex}}
%\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
\end{minipage}
\\
\hline
\end{tabular}
\caption{C Program, CFG and Abstract Conflict Graph}
\label{example}
\end{figure}
Let us consider a simple program in Figure~\ref{example} that checks 
the equivalence of a state-machine based interger multiplication against  
a specification that uses multiplication operator (*).  The assertion 
simply compares the result of two multiplication for an integer input $x$.  
To avoid overflow error, we assume $x$ has an upper bound $N$.  The 
control flow graph is shown in the middle column of Figure~\ref{example}.

We analyze the program using three different analyzers -- {\em bounded 
model checking (BMC)}, {\em abstract interpretation} and {\em ACDCL}.    
For $N=2$ and unwind limit of $bound=4$ for the {\em while} loop,
table~\ref{solver} shows the statistics from MiniSAT~\cite{minisat} solver 
for analysis using BMC.  The SAT solver makes 36 decisions, 1365 boolean 
constraint propagations and performs 5 conflict analysis with 65 conflict 
literals.  Moreover, the solver performs a restart, presumably to recover 
from bad decisions.

On the other hand, abstract interpretation using forward analysis 
is too imprecise to verify the safety of the above program using 
interval domain.  The imprecision is due to control-flow join at 
node $n1$.  State-of-the-art abstract interpretation tool, Astr{\'e}e, 
requires external hints, provided by manually annotating the code 
with partition directives at $n1$, to prove safety.  The partition 
directive helps the tool to analyze each program execution paths 
separately.  To do so, Astr{\'e}e requires a product of trace 
partitioning domain and an interval domain to determine safety.  
Consequently, the analysis becomes very expensive for a large value 
of $N$.  In general, the imprecision is either intended by the tool 
because such high precision analysis is normally not required for 
runtime error analysis or the imprecision is unavoidable due to the 
complexity of the application under analysis.  
%
\begin{table}
\begin{center}
{
\begin{tabular}{l|l|l|l|l|l}
\hline
Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
SAT & 36 & 1365 & 5 & 65 & 1 \\ \hline
ACDCL & 1 & 23 & 1 & 1 & 0 \\
\hline
\end{tabular}
}
\end{center}
\caption{Solver-statistics}
\label{solver}
\end{table}
%
The right-side of Figure~\ref{example} shows an analysis using ACDCL with 
interval domain.  The analysis associates an interval for each control 
location and variable, as shown in the abstract conflict graph.  Interval 
constraints generated from forward analysis in the initial deduction phase 
infers that $\langle x:[-\infty,2], spec:[2,4], impl:[-1,\infty],
state:[1,\infty], y:[1,+\infty] \rangle$.  Clearly, ACDCL can not prove safety.

Hence, ACDCL makes an arbitrary decision $x:[-\infty,1]$ to refine the analysis.  
The decision constrains the interval of $x$.  The deductions made during fixed point
iteration are represented by abstract conflict graph in right-side of 
Figure~\ref{fig:example}.   Nodes in the graph denotes the corresponding 
updates to the program variables in the cfg.  Interval analysis concludes 
that the $(n12:\bot)$, that is the decision $x:[-\infty,1]$ leads to a 
{\em conflict}. Intuitively, it means that the program is {\em safe} for 
$x:[-\infty,1]$.

A clause learning SAT solver would learn the reason for conflict at this 
point and then backtrack to a level such that the learnt clause is UNIT.  
Similar to conflict analysis phase in SAT solvers, ACDCL learns that 
(\(I: x = [2,+\infty]\)), that is all error traces must satisfy $(x \geq 2)$ 
at node $I$.  The analysis discards all interval constraints that
lead to the conflict and backtrack to decision level 0.  ACDCL then 
performs interval analysis with the learnt constraint $(x>=2)$ which again 
leads to a conflict.  The analysis cannot backtrack further and therefore 
it terminates proving that the program is safe.  A DPLL-style solver would 
perform proof by cases.  However, decision and clause learning are used 
to avoid case based reasoning and prevents enumeration behavior.

The result of the analysis using ACDCL is shown in Figure~\ref{example}.  
Compared to a SAT solver, ACDCL makes only a single decision which lead 
to 240 interval constraint propagations, and learns a single conflict clause 
involving only one literal.  Compared to abstract interpretation, ACDCL does 
not require external hints to proof the program, thus automatically performing 
program and property driven trace partitioning to generate proofs using decision 
and clause learning.
